Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-05-14 paper.tex
updated figures

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
 \includegraphics[width=0.49\columnwidth]{figures/flux.BMcompI.v11.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJ.v11.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompK.v11.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJs.v11.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\caption{\it Comparing the differential flux from different models: we
  show the continuum emission from DM benchmark (BM) models (blue
  dashed), electrons and positrons from DM BM models that inverse
  Compton upscattered both CMB photons (green dotted) as well as dust
  and star photons (orange dash-dotted), and CR induced gamma-ray
  emission (red solid). Each panel is associated with an individual DM
  BM model; upper left $\Ip$, lower left $\Kp$, upper right $\Jp$, and
  lower right $\Js$. The emission is calculated for the Fornax cluster
  using a point spread function of $0.1\deg$. The substructures boost
  the gamma-ray flux from IC upscattered CMB and continuum emission by
  570 while the IC upscattered SD photons is boosted by 20. Upper limits
from Fermi-LAT after 18 months 
 \protect \cite{2010ApJ...717L..71A} are also shown. We find
  bright prospects for detecting the continuum emission, that is
  dominating the total emission in the GeV energy range, in the models
  $\Ip$ and $\Kp$.}
 \label{fig:diff_BM}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.v11.0.1deg.1.6T.SubMass.elmu.SF300.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.BMv11.0.1deg.SubMass.noMW.woGal.eps}
\caption{\it Comparing the flux from different inverse Compton
  upscattered radiation fields. We show the differential inverse
  Compton emission induced by leptophilic DM in the left panel and by
  the $\Kp$ benchmark model in the right panel. The contribution from
  each individual radiation field from top line to bottom line; CMB
  (green dashed-triple-dotted), dust (purple dashed), stars (blue
  dashed-dotted). The sum of the three components are shown with the
  orange solid line.  The red solid line show the CR induced gamma-ray
  flux. The black arrow shows the differential upper limit from Fermi
  \protect \cite{2010ApJ...717L..71A} indicating that the model
  assumptions such as Sommerfeld and/or substructure boost are in
  conflict with the upper limit. We also show projected CTA point
  source sensitivities ($5\sigma$, 50h).  All fluxes are calculated
  for the Fornax cluster within $\rvir$ using a point spread function
  of $0.1\deg$. For this cluster, the boost from Sommerfeld is about
  130 and the enhancement due to substructures from IC upscattered CMB
  and SD photons is 570 and 20, respectively.}
 \label{fig:IR_comp}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
 \includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Fornax.v11.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Coma.v11.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Virgo.v11.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Perseus.v11.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\caption{\it Comparing the gamma-ray emission from different
  clusters. We show the differential flux from four clusters; Fornax
  (upper left), Coma (upper right), Virgo (lower left), and Perseus
  (lower right). We show CR induced emission (red solid), leptophilic
  (LP) models include both final state radiation and IC upscattered
  CMB, dust and starlight (green dash-dotted), and benchmark $\Kp$
  models (BM) include continuum emission, and IC upscattered CMB, dust
  and starlight (blue dashed). The arrows show the differential upper
  limits set by Fermi-LAT in the energy ranges $0.2-1$~GeV,
  $1-10$~GeV, and $10-100$~GeV from left to right. We show the
  individual components for the BM model and LP model for Virgo and
  Perseus, respectively. The flux from the clusters is integrated out
  to $\rvir$ using a point spread function of $0.1\deg$ and includes
  the boost from substructures. We find that the lower GeV energy
  regime is most constraining due to the peak sensitivity of Fermi at
  these energies and use this regime to get upper limits on boost
  factors.}
 \label{fig:clu_comp}
\end{minipage}
\end{figure*}

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/LP.const.diff.v11.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
 \caption{\it Constraining boost factors using flux upper limits. The
   leptophilic (LP) gamma-ray emission is derived within $\rvir$ using
   a point spread function of $0.1\deg$. In order not to overproduce
   the Fermi-LAT differential flux upper limits in the energy interval
   $1-10$~GeV the boost from substructures and Sommerfeld enhancement
   (SFE) is constrained for four clusters; Fornax (red), Virgo
   (orange), M49 (green), and NGC4636 (blue). The SFE has been
   rescaled from the MW value of 300 (c.f. Eq.~\ref{eq:B_sfe}) that is
   required to explain the electron and positron excess observed at
   Earth with LP DM (black arrow). If the DM interpretation is
   correct, we can constrain the smallest size of halos to be larger
   than $100\,\msun$. Contrarily, if the smallest size of DM halos is
   $10^{-6}\,\msun$, we can constrain the SFE to about 2 in Fornax,
   corresponding to a maximum SFE in the MW of about 5 which would be
   too small to support the DM annihilation hypothesis for the
   Pamela/Fermi excess.}
 \label{fig:boost_const}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=0.99\columnwidth]{figures/SB.v11.1GeV.SF300.SubMass.elmu.eps}
 \caption{\it Comparing the intrinsic surface brightness from
   different clusters without taking PSF effects into account. We show
   the CR induced emission (red), leptophilic emission (green), and
   emission from the $\Kp$ benchmark model (blue). The different
   line styles each represent a cluster; Fornax (solid), Coma (dotted),
   Virgo (dashed), and Perseus (dash-dotted). The boost from
   Sommerfeld effect and substructures is about 130 and 570,
   respectively. Notice the very similar shape and normalization of
   the DM models, while the CRs have a much larger scatter in the
   profiles due to the large variation of the gas fraction in these clusters.}
 \label{fig:SB_clu}
\end{figure}

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/SB.resolved.v11.1GeV.SF300.noSuB.vs.SubMass.elmu.eps}
 \caption{\it The impact of substructures on the intrinsic surface
   brightness profile of the Fornax cluster at 1~GeV. The emission
   induced by CRs is denoted by the red solid line, the leptophilic
   models by green lines, and the DM $\Kp$ benchmark model by blue
   lines. The dashed line show the brightness where the boost from
   substructures is included while the substructures are excluded in
   the dotted line. The boost from Sommerfeld and substructures is
   about 130 and 570, respectively. Notice the nearly flat profiles
   when substructures are included, and the relative large boost at
   $0.01\rvir$ that is due to projection of peripheral substructures
   onto the cluster center.}
 \label{fig:SB_sub}
\end{figure}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{figures/SB.v11.1GeV.SF300.noSuB.elmu.eps}
  \includegraphics[width=0.49\columnwidth]{figures/SB.fornax.v11.1GeV.SF300.noSuB.elmu.eps}
  \caption{\it The intrinsic surface brightness profiles without
    substructures at energies $E>1\ GeV$. We show the CR induced
    emission (red), leptophilic emission (green), and emission from
    the $\Kp$ benchmark model (blue).  LEFT PANEL: comparison of the surface
    brightness of different clusters; Fornax (solid), Coma (dotted),
    Virgo (dashed), and Perseus (dash-dotted). RIGHT PANEL: comparison of
    different emission components in Fornax; dotted lines show the
    inverse Compton (IC) upscattered CMB photons, dashed lines show
    the IC upscattered photons from stars and dust, and blue solid
    line show the continuum emission from the $\Kp$ benchmark model.}
 \label{fig:SB_clu_nosub}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{figures/SB.Fornax.v11.SF300.SubMass.elmu.eps}
  \includegraphics[width=0.49\columnwidth]{figures/SB.Virgo.v11.SF300.SubMass.elmu.eps}
\caption{\it Surface brightness predicted for Cherenkov telescopes at
  high energies. We show the emission above $100$~GeV and include the
  boost from substructures. We use a point spread function of
  $\psf=0.1\deg$ that is typical for Cherenkov
  telescopes as well as the Fermi-LAT at this energy. Left panel shows
  the Fornax cluster and right panel the Virgo cluster. The gamma-ray
  emission is derived for the following components; CRs (red solid),
  continuum emission from the DM $\Kp$ benchmark model (blue
  dash-dotted), as well as final state radiation (green dashed) and
  inverse Compton upscattered dust and starlight (green dotted) from
  leptophilic DM.}
 \label{fig:SB_IACTs}
\end{minipage}
\end{figure*}

>>>>

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
 \includegraphics[width=0.49\columnwidth]{figures/flux.BMcompI.v12.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJ.v12.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompK.v12.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJs.v12.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\caption{\it Comparing the differential flux from different models: we
  show the continuum emission from DM benchmark (BM) models (blue
  dashed), electrons and positrons from DM BM models that inverse
  Compton upscattered both CMB photons (green dotted) as well as dust
  and star photons (orange dash-dotted), and CR induced gamma-ray
  emission (red solid). Each panel is associated with an individual DM
  BM model; upper left $\Ip$, lower left $\Kp$, upper right $\Jp$, and
  lower right $\Js$. The emission is calculated for the Fornax cluster
  using a point spread function of $0.1\deg$. The substructures boost
  the gamma-ray flux from IC upscattered CMB and continuum emission by
  570 while the IC upscattered SD photons is boosted by 20. Upper limits
from Fermi-LAT after 18 months 
 \protect \cite{2010ApJ...717L..71A} are also shown. We find
  bright prospects for detecting the continuum emission, that is
  dominating the total emission in the GeV energy range, in the models
  $\Ip$ and $\Kp$.}
 \label{fig:diff_BM}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.v12.0.1deg.1.6T.SubMass.elmu.SF300.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.BMv12.0.1deg.SubMass.noMW.woGal.eps}
\caption{\it Comparing the flux from different inverse Compton
  upscattered radiation fields. We show the differential inverse
  Compton emission induced by leptophilic DM in the left panel and by
  the $\Kp$ benchmark model in the right panel. The contribution from
  each individual radiation field from top line to bottom line; CMB
  (green dashed-triple-dotted), dust (purple dashed), stars (blue
  dashed-dotted). The sum of the three components are shown with the
  orange solid line.  The red solid line show the CR induced gamma-ray
  flux. The black arrow shows the differential upper limit from Fermi
  \protect \cite{2010ApJ...717L..71A} indicating that the model
  assumptions such as Sommerfeld and/or substructure boost are in
  conflict with the upper limit. We also show projected CTA point
  source sensitivities ($5\sigma$, 50h).  All fluxes are calculated
  for the Fornax cluster within $\rvir$ using a point spread function
  of $0.1\deg$. For this cluster, the boost from Sommerfeld is about
  130 and the enhancement due to substructures from IC upscattered CMB
  and SD photons is 570 and 20, respectively.}
 \label{fig:IR_comp}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
 \includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Fornax.v12.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Coma.v12.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Virgo.v12.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Perseus.v12.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\caption{\it Comparing the gamma-ray emission from different
  clusters. We show the differential flux from four clusters; Fornax
  (upper left), Coma (upper right), Virgo (lower left), and Perseus
  (lower right). We show CR induced emission (red solid), leptophilic
  (LP) models include both final state radiation and IC upscattered
  CMB, dust and starlight (green dash-dotted), and benchmark $\Kp$
  models (BM) include continuum emission, and IC upscattered CMB, dust
  and starlight (blue dashed). The arrows show the differential upper
  limits set by Fermi-LAT in the energy ranges $0.2-1$~GeV,
  $1-10$~GeV, and $10-100$~GeV from left to right. We show the
  individual components for the BM model and LP model for Virgo and
  Perseus, respectively. The flux from the clusters is integrated out
  to $\rvir$ using a point spread function of $0.1\deg$ and includes
  the boost from substructures. We find that the lower GeV energy
  regime is most constraining due to the peak sensitivity of Fermi at
  these energies and use this regime to get upper limits on boost
  factors.}
 \label{fig:clu_comp}
\end{minipage}
\end{figure*}

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/LP.const.diff.v12.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
 \caption{\it Constraining boost factors using flux upper limits. The
   leptophilic (LP) gamma-ray emission is derived within $\rvir$ using
   a point spread function of $0.1\deg$. In order not to overproduce
   the Fermi-LAT differential flux upper limits in the energy interval
   $1-10$~GeV the boost from substructures and Sommerfeld enhancement
   (SFE) is constrained for four clusters; Fornax (red), Virgo
   (orange), M49 (green), and NGC4636 (blue). The SFE has been
   rescaled from the MW value of 300 (c.f. Eq.~\ref{eq:B_sfe}) that is
   required to explain the electron and positron excess observed at
   Earth with LP DM (black arrow). If the DM interpretation is
   correct, we can constrain the smallest size of halos to be larger
   than $100\,\msun$. Contrarily, if the smallest size of DM halos is
   $10^{-6}\,\msun$, we can constrain the SFE to about 2 in Fornax,
   corresponding to a maximum SFE in the MW of about 5 which would be
   too small to support the DM annihilation hypothesis for the
   Pamela/Fermi excess.}
 \label{fig:boost_const}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=0.99\columnwidth]{figures/SB.v12.1GeV.SF300.SubMass.elmu.eps}
 \caption{\it Comparing the intrinsic surface brightness from
   different clusters without taking PSF effects into account. We show
   the CR induced emission (red), leptophilic emission (green), and
   emission from the $\Kp$ benchmark model (blue). The different
   line styles each represent a cluster; Fornax (solid), Coma (dotted),
   Virgo (dashed), and Perseus (dash-dotted). The boost from
   Sommerfeld effect and substructures is about 130 and 570,
   respectively. Notice the very similar shape and normalization of
   the DM models, while the CRs have a much larger scatter in the
   profiles due to the large variation of the gas fraction in these clusters.}
 \label{fig:SB_clu}
\end{figure}

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/SB.resolved.v12.1GeV.SF300.noSuB.vs.SubMass.elmu.eps}
 \caption{\it The impact of substructures on the intrinsic surface
   brightness profile of the Fornax cluster at 1~GeV. The emission
   induced by CRs is denoted by the red solid line, the leptophilic
   models by green lines, and the DM $\Kp$ benchmark model by blue
   lines. The dashed line show the brightness where the boost from
   substructures is included while the substructures are excluded in
   the dotted line. The boost from Sommerfeld and substructures is
   about 130 and 570, respectively. Notice the nearly flat profiles
   when substructures are included, and the relative large boost at
   $0.01\rvir$ that is due to projection of peripheral substructures
   onto the cluster center.}
 \label{fig:SB_sub}
\end{figure}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{figures/SB.v12.1GeV.SF300.noSuB.elmu.eps}
  \includegraphics[width=0.49\columnwidth]{figures/SB.fornax.v12.1GeV.SF300.noSuB.elmu.eps}
  \caption{\it The intrinsic surface brightness profiles without
    substructures at energies $E>1\ GeV$. We show the CR induced
    emission (red), leptophilic emission (green), and emission from
    the $\Kp$ benchmark model (blue).  LEFT PANEL: comparison of the surface
    brightness of different clusters; Fornax (solid), Coma (dotted),
    Virgo (dashed), and Perseus (dash-dotted). RIGHT PANEL: comparison of
    different emission components in Fornax; dotted lines show the
    inverse Compton (IC) upscattered CMB photons, dashed lines show
    the IC upscattered photons from stars and dust, and blue solid
    line show the continuum emission from the $\Kp$ benchmark model.}
 \label{fig:SB_clu_nosub}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{figures/SB.Fornax.v12.SF300.SubMass.elmu.eps}
  \includegraphics[width=0.49\columnwidth]{figures/SB.Virgo.v12.SF300.SubMass.elmu.eps}
\caption{\it Surface brightness predicted for Cherenkov telescopes at
  high energies. We show the emission above $100$~GeV and include the
  boost from substructures. We use a point spread function of
  $\psf=0.1\deg$ that is typical for Cherenkov
  telescopes as well as the Fermi-LAT at this energy. Left panel shows
  the Fornax cluster and right panel the Virgo cluster. The gamma-ray
  emission is derived for the following components; CRs (red solid),
  continuum emission from the DM $\Kp$ benchmark model (blue
  dash-dotted), as well as final state radiation (green dashed) and
  inverse Compton upscattered dust and starlight (green dotted) from
  leptophilic DM.}
 \label{fig:SB_IACTs}
\end{minipage}
\end{figure*}
